---
title: "Crystal Lake Perch Exploratory Analysis"
author: "Noah Lottig"
date: "2/3/2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#get data
#lter_fish <- read_csv("https://lter.limnology.wisc.edu/file/12848/download?token=-0G3eTcJCnat50PvBQYK2hTp9d8BWIA447_iz0NcH94")
```

## Crystal Lake Perch 

Preliminary analysis of long-term LTER data in Crystal Lake examining patterns in Perch and Smelt following the whole-lake mixing project. In late July 2017 NRL and LTER fish crew observed that perch CPUE appeared to be higher than ever recalled (NRL involved in fish crew since 2009).

```{r perch}
CR_Fish = lter_fish %>% filter(lakeid=="CR") %>% filter(spname =="YELLOWPERCH" | spname == "RAINBOWSMELT")
CR_Fish$CPUE = round(CR_Fish$total_caught/CR_Fish$effort,digits=3)
CR_Fish$spname = as.factor(CR_Fish$spname)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
